### 2022-05-24 Manuscript/Manuscript.tex
Update appendix 2 references

\textit{\textbf{Example}: See lines 2 - 11 in the \nameref{S2_Appendix} for the corresponding JSON..}

@@

\textit{\textbf{Example}: Pichu, Pikachu and Raichu are all species within the Pikachuidae. They each have a default and a ``Shiny'' morph, which is to be determined. The default Pikachu morph also has a subdivision into the sexes, but as the morph is tagged as an endpoint the user will not be asked further questions to determine these once the morph has been determined. See lines 27 - 111 \nameref{S2_Appendix} for the corresponding JSON, and Fig.~\ref{fig3} for a graphical representation.}

@@

% Image?

\textit{\textbf{Example}: Pikachu is always an electric type. See lines 287 - 292 in the \nameref{S2_Appendix} for the corresponding JSON.}

% Image?

\textit{\textbf{Example}: Pikachu is never blue. Note that it is not inferred that Pikachus therefore are red. See lines 293 - 299 in the \nameref{S2_Appendix} for the corresponding JSON.}

% Image?

\textit{\textbf{Example}: Pikachu has a double-lobed tail in 50\% of the cases. See lines 300 - 313 in the \nameref{S2_Appendix} for the corresponding JSON.}
\subsection*{
Multilingualism
}
Rather than referring to a single language by its ISO 639-1 code, one can make a key multilingual by specifying an array of language codes that are supported. When doing so, strings within the key that differ between languages have to be given as localizedStrings, objects containing a version of each of the supported languages. To support different script types, also strings like people's names support localizedStrings.

@@

\textit{\textbf{Example}: A creator name that needs no translation, and a character that does, within a multilingual English/Dutch key}
\begin{verbatim}
"language": ["en", "nl"],
"creator": "person:wouterkoch",
"persons": [
   {
     "id": "person:wouterkoch",
     "name": "Wouter Koch"
   }
],
"characters": [
   {
     "id": "character:eye_color",
     "title": {
        "en": "Color of the eyes",
        "nl": "Kleur van de ogen"
     },
     "states": [
       {
         "id": "state:blue_eyes",
         "title": {"en": "Blue", "nl": "Blauw"}
       },
       {
         "id": "state:red_eyes",
         "title": {"en": "Red", "nl": "Rood"}
       }
     ]
   }
]

@@

\textit{\textbf{Example}: An organization with names and urls within a multilingual Norwegian/English key}
\begin{verbatim}
"language": ["no", "en"],
"publisher": "organization:ntnu",
"organizations": [
   {
     "id": "organization:ntnu",
     "name": {
       "no": "Norges teknisk-naturvitenskapelige universitet",
       "en": "Norwegian University of Science and Technology"
     },
     "url": {
       "no": "https://www.ntnu.no",
       "en": "https://www.ntnu.edu"
     }
   }
]

@@

Persons and institutions can have resources such as urls and mediaFiles connected to them, e.g. institutional websites, portraits and logos.

% Image?

\textit{\textbf{Example}: Definition of a person and an organization. See lines 12 - 25 in the \nameref{S2_Appendix} for the corresponding JSON.}
\subsection*{
Geographic and taxonomic scope
}
Regions can be defined using a GeoJSON MultiPolygon. On the top level, ``geography'' can be specified to indicate what geographic region the key covers. One can also define a geography at the taxon or statement level, to provide information on where the taxon occurs, and in which region a taxon has that particular relationship to a character. Within a region specified as the geography of a statement, the statement with a geography takes precedence over the conflicting statement without a stated geography.

% Image?

\textit{\textbf{Example}: Kangaskhan only occurs in Australia. See lines 120 - 169 in the \nameref{S2_Appendix} for the corresponding JSON.}

% Image?

\textit{\textbf{Example}: The rain form of Castform has a higher frequency in the rainy city of Bergen. See lines 325 - 368 in the \nameref{S2_Appendix} for the corresponding JSON.}
\subsection*{
Logical premises
}
Not all characters make sense in every context, and it is sometimes desirable to be able to specify conditions that must be met before a character is shown to the user. Such a condition may be that the user has given a certain answer to some other character in the key first. To this end, characters can have a logical premise, a JavaScript notation string referring to which facts have to be established for it to be relevant. Logical premises can also relate several facts combined, including numerical values, and can be built using unary and binary operators in JavaScript notation.

@@

% Image?

\textit{\textbf{Example}: Pikachu caught during special events have hats, in contrast to those caught outside of such events. Other Pokémon (e.g. Honchkrow) always have hats. This allows for the characteristics of the hat to be used in identification, but only when one is not only left solely with Pokémon that can have hats, but when it has been established that the target Pokémon in fact is wearing one. See lines 257 - 271 in the \nameref{S2_Appendix} for the corresponding JSON.}
\subsection*{
Numerical values
}
The possible values that a character can have can take the form of a set of discrete states, such as ``present''/``absent'', or ``red''/``blue''. Others may however come in the form of numerical values, such as counts or measurements. In a key this can be implemented with a character of the type ``numerical''. One can specify the range (min and max) the character can have, the step size and the unit. Statements specifying values for such characters to any taxon can have a range, or a single value. In cases where more advanced metrics are needed, such as a probability function over the numerical values for use in Bayesian analysis, this can be implemented through a call to an external service (see below).

% Image?

\textit{\textbf{Example}: The weight of a Pokémon varies, both between individuals within a species and between species. To distinguish Pokémon based on their weight, this is added as a numerical character, where weights can be specified in whole kg within a given range. A statement can then be included to register the possible weight range of a given taxon. See lines 272 - 284 (character) and lines 314 - 324 (statement) in \nameref{S2_Appendix} for the corresponding JSON.}
\subsection*{
External services
}
Various units in a key may be subject to changes that are ideally managed outside the key, in designated centralized systems. Taxon names can change, as can urls to supplemental information, media files, geographic ranges etc. Other parameters, such as the probability distribution of a numerical character for a given taxon, may be too detailed for direct inclusion in a key. It is best practice to not duplicate such resources, but to harvest these via an API or other service interface. To facilitate this, Clavis allows the specification of external services, where documentation for use of the service can be linked. Various units in the key, such as taxa, can refer to such external services through one or more externalResources, connecting the service and the relevant external id to the taxon at hand.

@@

% Image?

\textit{\textbf{Example}: To allow for features such as the retrieval of the name of a Pokémon in different languages, one can use a Wikidata query. See lines 458 - 462 (service) and lines 52 - 55 (externalResource) in the \nameref{S2_Appendix} for the corresponding JSON.}

% Image?

@@

% Image?

\textit{\textbf{Example}: the weight of a Pokémon can be a useful characteristic as their weight ranges vary. One can only weigh a Pokémon once it has been caught, however. The user needs to be aware of this so that in situations where a Pokémon cannot be caught (e.g. when identifying from a picture or being out of Pokéballs) such questions can be filtered out. One can also add a guide to how one catches a Pokémon. See lines 272 - 284 (character) and lines 441 - 455 (userRequirement) in the \nameref{S2_Appendix} for the corresponding JSON.}

@@

% Image?

\textit{\textbf{Example}: The states in the tail character of Pikachu refer to images illustrating the different tail shapes. See lines 227 - 240 (states) and lines 465 - 511 (mediaElements) in the \nameref{S2_Appendix} for the corresponding JSON.}
\subsection*{
Descriptions
}
There are many aspects of an identification key where a more elaborate description is desirable or even required. To this end, many elements can have a short and/or extended description in valid markdown, as well as a url to an online description.

% Image?

\textit{\textbf{Example}: A description can be used to explain to the user how to catch a Pokémon where relevant, by including a short description and a link to a page with more information. See lines 446 - 455 in the \nameref{S2_Appendix} for the corresponding JSON.}
\subsection*{
Followup
}
Once a key has identified a taxon as far as it is intended to, i.e. when it has narrowed it down to an end point taxon, the user is presented with the result. It may however be possible to further determine the result. If a key exists somewhere that can help with this, it can be referred to as a followup key from the relevant taxon. It can be either a url, or a reference to an (external) service. This feature can also be used to split keys into several smaller keys that refer to one another through this mechanism.

% Image?

@@

\textit{\textbf{Example}: Pikachu contains the colors yellow, black and red. The closely related Raichu contains yellow, brown, white and red. Setting the color character to be of non-exclusive type allows the user to indicate all the colors that each of them has. See lines 200 - 224 in the \nameref{S2_Appendix} for the corresponding JSON, and Fig.~\ref{fig4} for a graphical representation.}

\begin{figure}[!h]
  \caption{\textit{\textbf{The colors of Pikachu and Raichu}:
Pikachu contains the colors yellow, black and red. The closely related Raichu contains yellow, brown, white and red. Setting the color character to be of non-exclusive type allows the user to indicate all the colors that each of them has.
}}
 \label{fig4}
\end{figure}

>>>>

\textit{\textbf{Example}: See lines 2 - 10 in the \nameref{S2_Appendix} for the corresponding JSON..}

@@

\textit{\textbf{Example}: Pichu, Pikachu and Raichu are all species within the Pikachuidae. They each have a default and a ``Shiny'' morph, which is to be determined. The default Pikachu morph also has a subdivision into the sexes, but as the morph is tagged as an endpoint the user will not be asked further questions to determine these once the morph has been determined. See lines 25 - 94 in the \nameref{S2_Appendix} for the corresponding JSON, and Fig.~\ref{fig3} for a graphical representation.}

@@

\textit{\textbf{Example}: Pikachu is always an electric type. See lines 264 - 270 in the \nameref{S2_Appendix} for the corresponding JSON.}

\textit{\textbf{Example}: Pikachu is never blue. Note that it is not inferred that Pikachus therefore are red. See lines 271 - 277 in the \nameref{S2_Appendix} for the corresponding JSON.}

\textit{\textbf{Example}: Pikachu has a double-lobed tail in 50\% of the cases. See lines 278 - 291 in the \nameref{S2_Appendix} for the corresponding JSON.}
\subsection*{
Multilingualism
}
Rather than referring to a single language by its ISO 639-1 code, one can make a key multilingual by specifying an array of language codes that are supported. When doing so, strings within the key that differ between languages have to be given as localizedStrings, objects containing a version of each of the supported languages. To support different script types, also strings like people's names support localizedStrings.

@@

\textit{\textbf{Example}: A creator name that needs no translation, and a character that does, within a multilingual English/Dutch key}

\begin{verbatim}
"language": ["en", "nl"],
"creator": "person:wouterkoch",
"persons": [
   {
     "id": "person:wouterkoch",
     "name": "Wouter Koch"
   }
],
"characters": [
   {
     "id": "character:eye_color",
     "title": {
        "en": "Color of the eyes",
        "nl": "Kleur van de ogen"
     },
     "states": [
       {
         "id": "state:blue_eyes",
         "title": {"en": "Blue", "nl": "Blauw"}
       },
       {
         "id": "state:red_eyes",
         "title": {"en": "Red", "nl": "Rood"}
       }
     ]
   }
]

@@

\textit{\textbf{Example}: An organization with names and urls within a multilingual Norwegian/English key}

\begin{verbatim}
"language": ["no", "en"],
"publisher": "organization:ntnu",
"organizations": [
   {
     "id": "organization:ntnu",
     "name": {
       "no": "Norges teknisk-naturvitenskapelige universitet",
       "en": "Norwegian University of Science and Technology"
     },
     "url": {
       "no": "https://www.ntnu.no",
       "en": "https://www.ntnu.edu"
     }
   }
]

@@

Persons and institutions can have resources such as urls and mediaFiles connected to them, e.g. institutional websites, portraits and logos.
\textit{\textbf{Example}: Definition of a person and an organization. See lines 11 - 23 in the \nameref{S2_Appendix} for the corresponding JSON.}
\subsection*{
Geographic and taxonomic scope
}
Regions can be defined using a GeoJSON MultiPolygon. On the top level, ``geography'' can be specified to indicate what geographic region the key covers. One can also define a geography at the taxon or statement level, to provide information on where the taxon occurs, and in which region a taxon has that particular relationship to a character. Within a region specified as the geography of a statement, the statement with a geography takes precedence over the conflicting statement without a stated geography.

\textit{\textbf{Example}: Kangaskhan only occurs in Australia. See lines 95 - 146 in the \nameref{S2_Appendix} for the corresponding JSON.}

\textit{\textbf{Example}: The rain form of Castform has a higher frequency in the rainy city of Bergen. See lines 302 - 347 in the \nameref{S2_Appendix} for the corresponding JSON.}
\subsection*{
Logical premises
}
Not all characters make sense in every context, and it is sometimes desirable to be able to specify conditions that must be met before a character is shown to the user. Such a condition may be that the user has given a certain answer to some other character in the key first. To this end, characters can have a logical premise, a JavaScript notation string referring to which facts have to be established for it to be relevant. Logical premises can also relate several facts combined, including numerical values, and can be built using unary and binary operators in JavaScript notation.

@@

\textit{\textbf{Example}: Pikachu caught during special events have hats, in contrast to those caught outside of such events. Other Pokémon (e.g. Honchkrow) always have hats. This allows for the characteristics of the hat to be used in identification, but only when one is not only left solely with Pokémon that can have hats, but when it has been established that the target Pokémon in fact is wearing one. See lines 220 - 248 in the \nameref{S2_Appendix} for the corresponding JSON.}
\subsection*{
Numerical values
}
The possible values that a character can have can take the form of a set of discrete states, such as ``present''/``absent'', or ``red''/``blue''. Others may however come in the form of numerical values, such as counts or measurements. In a key this can be implemented with a character of the type ``numerical''. One can specify the range (min and max) the character can have, the step size and the unit. Statements specifying values for such characters to any taxon can have a range, or a single value. In cases where more advanced metrics are needed, such as a probability function over the numerical values for use in Bayesian analysis, this can be implemented through a call to an external service (see below).

\textit{\textbf{Example}: The weight of a Pokémon varies, both between individuals within a species and between species. To distinguish Pokémon based on their weight, this is added as a numerical character, where weights can be specified in whole kg within a given range. A statement can then be included to register the possible weight range of a given taxon. See lines 249 - 261 (character) and lines 292 - 301 (statement) in the \nameref{S2_Appendix} for the corresponding JSON.}
\subsection*{
External services
}
Various units in a key may be subject to changes that are ideally managed outside the key, in designated centralized systems. Taxon names can change, as can urls to supplemental information, media files, geographic ranges etc. Other parameters, such as the probability distribution of a numerical character for a given taxon, may be too detailed for direct inclusion in a key. It is best practice to not duplicate such resources, but to harvest these via an API or other service interface. To facilitate this, Clavis allows the specification of external services, where documentation for use of the service can be linked. Various units in the key, such as taxa, can refer to such external services through one or more externalResources, connecting the service and the relevant external id to the taxon at hand.

@@

\textit{\textbf{Example}: To allow for features such as the retrieval of the name of a Pokémon in different languages, one can use a Wikidata query. See lines 435 - 439 (service) and lines 47 - 50 (externalResource) in the \nameref{S2_Appendix} for the corresponding JSON.}

@@

\textit{\textbf{Example}: the weight of a Pokémon can be a useful characteristic as their weight ranges vary. One can only weigh a Pokémon once it has been caught, however. The user needs to be aware of this so that in situations where a Pokémon cannot be caught (e.g. when identifying from a picture or being out of Pokéballs) such questions can be filtered out. One can also add a guide to how one catches a Pokémon. See lines 249 - 261 (character) and lines 420 - 432 (userRequirement) in the \nameref{S2_Appendix} for the corresponding JSON.}

@@

\textit{\textbf{Example}: The states in the tail character of Pikachu refer to images illustrating the different tail shapes. See lines 207 - 218 (states) and lines 449 - 498 (mediaElements) in the \nameref{S2_Appendix} for the corresponding JSON.}
\subsection*{
Descriptions
}
There are many aspects of an identification key where a more elaborate description is desirable or even required. To this end, many elements can have a short and/or extended description in valid markdown, as well as a url to an online description.

\textit{\textbf{Example}: A description can be used to explain to the user how to catch a Pokémon where relevant, by including a short description and a link to a page with more information. See lines 425 - 431 in the \nameref{S2_Appendix} for the corresponding JSON.}
\subsection*{
Followup
}
Once a key has identified a taxon as far as it is intended to, i.e. when it has narrowed it down to an end point taxon, the user is presented with the result. It may however be possible to further determine the result. If a key exists somewhere that can help with this, it can be referred to as a followup key from the relevant taxon. It can be either a url, or a reference to an (external) service. This feature can also be used to split keys into several smaller keys that refer to one another through this mechanism.

@@

\textit{\textbf{Example}: Pikachu contains the colors yellow, black and red. The closely related Raichu contains yellow, brown, white and red. Setting the color character to be of non-exclusive type allows the user to indicate all the colors that each of them has. See lines 177 - 203 in the \nameref{S2_Appendix} for the corresponding JSON, and Fig.~\ref{fig4} for a graphical representation.}

\begin{figure}[!h]
  \caption{\textit{\textbf{The colors of Pikachu and Raichu}:
By adding statements for the non-exclusive character for the color of Pikachu and Raichu, the user can pick and/or exclude freely from a list of colors in the identification key.
}}
 \label{fig4}
\end{figure}
